---
title: "Example Shot Charts"
author: "Sam Hughes"
date: "2023-03-30"
output: pdf_document
---


```{r}
library(tidyverse)
```

## Get data

Note: make sure data is in its own folder named "data"

```{r}
l = list()
for (season in c(2007:2022)) {
  l = l %>% append(list(readRDS(paste0('./data/shot_data_', season, '.rds'))))
}

df = bind_rows(l)

rm(l)
```

Data cleaning

```{r}
# Make sure 0 distance shots are from restricted area
df = df[df$SHOT_ZONE_BASIC == 'Restricted Area' | df$SHOT_DISTANCE > 0, ]

# Make sure 3 pointers fall under appropriate category
df = df[!(df$SHOT_TYPE == '3PT Field Goal' & 
            df$SHOT_ZONE_BASIC %in% c('In The Paint (Non-RA)', 'Mid-Range', 
                                      'Restricted Area')), ]

# Make sure 2 pointers fall under appropriate category
df = df[!(df$SHOT_TYPE == '2PT Field Goal' & 
            df$SHOT_ZONE_BASIC %in% c('Left Corner 3', 'Right Corner 3', 
                                     'Above the Break 3', 'Backcourt')), ]

# Filter out no shots
df = df[df$ACTION_TYPE != 'No Shot', ]
```



Restructure Action Types
```{r}
df$ACTION_TYPE = df$ACTION_TYPE %>% tolower()

# simplification
df$ACTION_TYPE = gsub('tip dunk', 'putback dunk', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('hook bank', 'hook', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('jump bank shot', 'jump shot', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('bank shot', 'jump shot', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('tip layup', 'tip', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('jump hook', 'hook', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('pull-up', 'driving', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('pullup', 'driving', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('running driving', 'driving', df$ACTION_TYPE)

# remove
df$ACTION_TYPE = gsub('bank ', '', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('slam ', '', df$ACTION_TYPE)
df$ACTION_TYPE = gsub('finger roll ', '', df$ACTION_TYPE)

df$SHOT_TECHNIQUE = word(df$ACTION_TYPE, -2, -1)
df[df$SHOT_TECHNIQUE == 'layup shot', 'SHOT_TECHNIQUE'] = 'layup'
df[df$SHOT_TECHNIQUE == 'dunk shot', 'SHOT_TECHNIQUE'] = 'dunk'
df[df$SHOT_TECHNIQUE == 'fadeaway shot', 'SHOT_TECHNIQUE'] = 'jump shot'

df$SHOT_TECHNIQUE = factor(df$SHOT_TECHNIQUE, 
                           levels=c('jump shot', 'dunk', 'layup', 'hook shot', 'tip shot'))

df$running_shot = if_else(grepl('running', df$ACTION_TYPE), 1, 0)
df$turnaround_shot = if_else(grepl('turnaround', df$ACTION_TYPE), 1, 0)
df$driving_shot = if_else(grepl('driving', df$ACTION_TYPE), 1, 0)
df$putback_shot = if_else(grepl('putback', df$ACTION_TYPE), 1, 0)
df$stepback_shot = if_else(grepl('step', df$ACTION_TYPE), 1, 0)
df$cutting_shot = if_else(grepl('cutting', df$ACTION_TYPE), 1, 0)
df$reverse_shot = if_else(grepl('reverse', df$ACTION_TYPE), 1, 0)
df$fadeaway_shot = if_else(grepl('fadeaway', df$ACTION_TYPE), 1, 0)
df$alleyoop_shot = if_else(grepl('alley', df$ACTION_TYPE), 1, 0)
df$floater_shot = if_else(grepl('floating', df$ACTION_TYPE), 1, 0)
```

## Example Shot Charts

### Draw Court Function
Citation for court dimensions: https://towardsdatascience.com/make-a-simple-nba-shot-chart-with-python-e5d70db45d0d
```{r}
draw_court = function(color='black', hoop_color='orange', background=NULL) {
  return(list(geom_segment(mapping=aes(x=-250, xend=250, y=-40, yend=-40), size=1, color=color), 
  geom_segment(mapping=aes(x=-250, xend=250, y=430, yend=430), size=1, color=color),
  geom_segment(mapping=aes(x=-250, xend=-250, y=-40, yend=430), size=1, color=color),
  geom_segment(mapping=aes(x=250, xend=250, y=-40, yend=430), size=1, color=color),
  # three point line
  geom_segment(mapping=aes(x=-220, xend=-220, y=-40, yend=90), size=1, color=color),
  geom_segment(mapping=aes(x=220, xend=220, y=-40, yend=90), size=1, color=color),
  geom_curve(mapping=aes(x=-220, xend=220, y=90, yend=90), size=1, 
             curvature=-0.78, lineend='round', color=color),
  # key
  geom_segment(mapping=aes(x=-80, xend=-80, y=-40, yend=150), size=1, color=color),
  geom_segment(mapping=aes(x=80, xend=80, y=-40, yend=150), size=1, color=color),
  geom_segment(mapping=aes(x=-60, xend=-60, y=-40, yend=150), size=1, color=color),
  geom_segment(mapping=aes(x=60, xend=60, y=-40, yend=150), size=1, color=color),
  geom_segment(mapping=aes(x=-80, xend=80, y=150, yend=150), size=1, color=color),
  geom_curve(mapping=aes(x=-60, xend=60, y=150, yend=150), size=1, 
             curvature=-1, lineend='round', color=color),
  # hoop
  geom_segment(mapping=aes(x=-30, xend=30, y=0, yend=0), size=1, color=color),
  geom_curve(mapping=aes(x=-10, xend=10, y=13, yend=13), size=1, 
             curvature=-1, color=hoop_color, lineend='round', color=color),
  geom_curve(mapping=aes(x=-10, xend=10, y=13, yend=13), size=1, 
             curvature=1, color=hoop_color, lineend='round', color=color),
  lims(x=c(-250, 250), y=c(-40, 430)),
  theme_minimal(), 
  theme(panel.grid=element_blank(), 
        panel.background=element_rect(fill=background, linetype='blank'), 
        plot.background=element_rect(fill=background, linetype='blank'), 
        axis.text=element_blank(), axis.title=element_blank(), 
        aspect.ratio=4/3.76)))
}
```



Note: to help with the fact that so many shots are taken within a very close region to the hoop, I am manually adjusting the scale of the frequency plots. In the hexbins, I adjust to a log scale, and in the heatmaps, this is the reason for the long 'breaks' formula. 

### Frequency with heat maps
```{r}
ggplot(data=df %>% filter(PLAYER_NAME == 'Marcus Smart'), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_density2d_filled(contour_var='ndensity', 
                        breaks=seq(0, 1, 0.05)^2*2/3+seq(0, 1, 0.05)*1/3, 
                        show.legend=FALSE) +
  draw_court(color='grey')
```


```{r}
ggplot(data=df %>% filter(PLAYER_NAME == 'Ray Allen'), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_density2d_filled(contour_var='ndensity', 
                        breaks=seq(0, 1, 0.05)^2*2/3+seq(0, 1, 0.05)*1/3, 
                        show.legend=FALSE) +
  draw_court(color='grey')
```


```{r}
ggplot(data=df %>% filter(PLAYER_NAME == 'Kevin Garnett'), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_density2d_filled(contour_var='ndensity', 
                        breaks=seq(0, 1, 0.05)^2*2/3+seq(0, 1, 0.05)*1/3, 
                        show.legend=FALSE) +
  draw_court(color='grey')
```





```{r}
ggplot(data=df %>% filter(PLAYER_NAME == 'Stephen Curry' & 
                            season == 2018), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_density2d_filled(contour_var='ndensity', 
                        breaks=seq(0, 1, 0.05)^2*2/3+seq(0, 1, 0.05)*1/3, 
                        show.legend=FALSE) +
  draw_court(color='grey')
```






### Frequency with hex bins

(This is the same plot as above except with hex bins)
```{r}
ggplot(data=df %>% filter(PLAYER_NAME == 'Stephen Curry' & 
                            season == 2018), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=FALSE) + 
  scale_fill_gradient(low='white', 
                      high='orange', 
                      trans='log') +
  labs(title='   Steph Curry') +
  draw_court(hoop_color='black')
```

Steph Curry and Ben Simmons are both listed as point guards, yet have vastly different shot charts. 
```{r}
ggplot(data=df %>% filter(PLAYER_NAME == 'Ben Simmons' & 
                            season == 2018), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=FALSE) + 
  scale_fill_gradient(low='white', 
                      high='purple', 
                      trans='log') +
  labs(title='   Ben Simmons') +
  draw_court()
```

Comparison of the same team from 2007 vs 2022
```{r}
ggplot(data=df %>% filter(TEAM_NAME == 'Boston Celtics' & 
                            season == 2007), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=FALSE) + 
  scale_fill_gradient(low='white', 
                      high='orange', 
                      trans='log') +
  draw_court(hoop_color='black')
```

```{r}
ggplot(data=df %>% filter(TEAM_NAME == 'Boston Celtics' & 
                            season == 2022), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=FALSE) + 
  scale_fill_gradient(low='white', 
                      high='orange', 
                      trans='log') +
  draw_court(hoop_color='black')
```

```{r}
set.seed(22)

a = ggplot(data=df %>% filter(season == 2007) %>% sample_n(5000), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=FALSE) + 
  scale_fill_steps(low='white', 
                      high='orange', 
                      trans='log') +
  draw_court(hoop_color='black') + 
  labs(title='   2007 Shot Frequencies')

a
```

```{r}
set.seed(22)

b = ggplot(data=df %>% filter(season == 2022) %>% sample_n(5000), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=FALSE) + 
  scale_fill_steps(low='white', 
                      high='orange', 
                      trans='log') +
  draw_court(hoop_color='black') +
  labs(title='   2022 Shot Frequencies')
```

```{r}
library(ggpubr)

plot = ggarrange(a, b, ncol=2)

annotate_figure(plot, top='Shot Frequencies')
```

```{r}
set.seed(22)

ggplot(data=df %>% filter(season %in% c(2007, 2022)) %>% 
         group_by(season) %>% sample_n(10000), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=TRUE) + 
  facet_wrap(~season, switch='x') +
  scale_fill_gradient(low='white', 
                      high='orange', 
                      trans='log', 
                      labels=c('', 'Less\nFrequent', '', 'More\nFrequent', '')) +
  draw_court(hoop_color='black') +
  labs(title='                                 Shot Frequencies in 2007 vs 2022', 
       fill=NULL) +
  theme(strip.text.x=element_text(size=12))
```

```{r}
df_2007_2022 = df %>% filter(season %in% c(2007, 2022)) %>% 
  group_by(season) %>% sample_n(20000)


# season_ = c(2007, 2022)
# LOC_X_ = seq(-230, 230, 10)
# LOC_Y_ = seq(0, 240, 10)
# temp = expand.grid(season_, LOC_X_, LOC_Y_)
# colnames(temp) = c('season', 'LOC_X', 'LOC_Y')
# temp$SHOT_DISTANCE = round((temp$LOC_X^2 + temp$LOC_Y^2)^0.5/10)
# temp = temp %>% filter(SHOT_DISTANCE <= 24)
# 
# df_2007_2022 = rbind(df_2007_2022, temp)
```




```{r}
set.seed(22)

ggplot(data=df_2007_2022, 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=TRUE) + 
  facet_wrap(~season, switch='x') +
  scale_fill_stepsn(colors=c('white', '#ffcbd1', '#ee6b6e', '#ff2c2c', '#c30010'), 
                    trans='log', 
                    labels=c('Less\nFrequent', '', '', 'More\nFrequent')) +
  draw_court(hoop_color='black') +
  labs(title='                                 Shot Frequencies in 2007 vs 2022', 
       fill=NULL) +
  theme(strip.text.x=element_text(size=12))
```





```{r}
ggplot(data=df %>% filter(TEAM_NAME == 'Golden State Warriors' & 
                            season == 2022), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=FALSE) + 
  scale_fill_gradient(low='white', 
                      high='orange', 
                      trans='log') +
  draw_court(hoop_color='black')
```

```{r}
ggplot(data=df %>% filter(TEAM_NAME == 'Los Angeles Lakers' & 
                            season == 2022), 
       mapping=aes(x=LOC_X, y=LOC_Y)) +
  geom_hex(show.legend=FALSE) + 
  scale_fill_gradient(low='white', 
                      high='orange', 
                      trans='log') +
  draw_court(hoop_color='black')
```




### Field goal percentage with hex bins 
This is the percentage of made shots over total shots within each bin

(These plots would probably be better with a model so it could display field goal percentage above expected)

```{r}
ggplot(data=df %>% filter(PLAYER_NAME == 'Damian Lillard' & 
                            season == 2022), 
       mapping=aes(x=LOC_X, y=LOC_Y, z=SHOT_MADE_FLAG)) +
  stat_summary_hex() +
  scale_fill_gradient(low='light blue', 
                      high='orange') +
  draw_court(hoop_color='black')
```








```{r}
df_shot_chart = df %>% filter(SHOT_DISTANCE < 28 & LOC_X %% 2 == 0 & LOC_Y %% 2 == 0) %>% 
  group_by(LOC_X, LOC_Y) %>% sample_n(1)

nrow(df_shot_chart)
```

```{r}
df_shot_chart$points = if_else(df_shot_chart$SHOT_TYPE == '3PT Field Goal', 
                               3*df_shot_chart$SHOT_MADE_FLAG, 2*df_shot_chart$SHOT_MADE_FLAG)

mean(df_shot_chart$points)
```

```{r}
a = ggplot(data=df_shot_chart, 
       mapping=aes(x=LOC_X, y=LOC_Y, z=SHOT_MADE_FLAG)) +
  stat_summary_hex() +
  scale_fill_gradient(low='light blue', 
                      high='orange') +
  draw_court(hoop_color='black') +
  labs(fill='Field Goal Percentage')

a
```



```{r}
b = ggplot(data=df_shot_chart, 
       mapping=aes(x=LOC_X, y=LOC_Y, z=points)) +
  stat_summary_hex(bins=30) +
  scale_fill_gradient(low='light blue', 
                      high='orange') +
  draw_court(hoop_color='black') +
  labs(fill='Average Points Scored')

b
```






